---
title: "AE 02: Life expectancy and healthcare expenditure"
date: "Jan 14, 2025"
editor: visual
html-math-method:
  method: mathjax
  url: "https://cdn.jsdelivr.net/npm/mathjax@3/es5/tex-mml-chtml.js"
bibliography: references.bib
---

::: callout-important
For this AE, you will discuss the questions in groups. This AE does not count towards the Application Exercise grade.
:::

```{r load-packages}
#| message: false
library(tidyverse)
library(readxl)
library(tidymodels)
library(patchwork)
```

## Introduction

The data set comes from @zarulli2021, who analyze the effects of a country's healthcare expenditures and other factors on the country's life expectancy. The data are originally from the [Human Development Database](http://hdr.undp.org/en/data) and [World Health Organization](https://apps.who.int/nha/database/).

This AE will focus on the following variables:

-   `life_exp`: *The average number of years that a newborn could expect to live, if he or she were to pass through life exposed to the sex- and age-specific death rates prevailing at the time of his or her birth, for a specific year, in a given country, territory, or geographic area.* ( from the [World Health Organization](https://www.who.int/data/gho/indicator-metadata-registry/imr-details/65#:~:text=Definition%3A,%2C%20territory%2C%20or%20geographic%20area.))
-   `health_pct_gdp`: *Spending on healthcare goods and services, expressed as a percentage of GDP. It excludes capital health expenditures such as buildings, machinery, information technology and stocks of vaccines for emergency or outbreaks.*

[Click here](https://doi.org/10.1371/journal.pone.0253450) for the original research paper and a full list of variables in the original data set.

```{r load-data}
#| message: false
life_exp <- read_excel("data/life-expectancy-data.xlsx") |> 
  rename(life_exp = `Life_expectancy_at_birth`, 
         health_pct_gdp = `Domestic_general_government_health_expenditure_pct_of_GDP`)
```

```{r}
life_exp |>
  select(life_exp, health_pct_gdp) |>
  glimpse()
```

## Exercises

We begin by visualizing the distributions of life expectancy, health expenditure percentage, and the relationship between these two variables.

```{r eda}
#| warning: false
#| message: false
#| fig-align: center

p1 <- ggplot(life_exp, aes(x = life_exp)) +
  geom_histogram() + 
  labs(x = "Life expectancy")

p2 <- ggplot(life_exp, aes(x = health_pct_gdp)) +
  geom_histogram() + 
  labs(x = "% Health expenditure")

p3 <- ggplot(life_exp, aes(x = health_pct_gdp, y = life_exp)) +
  geom_point() + 
  labs(x = "% Health expenditure", 
       y = "Life expectancy")

(p1 | p2) / p3
```

### Exercise 1

Describe the relationship between life expectancy and healthcare expenditure as a percentage of the GDP. Comment on how we expect the life expectancy to change as the percentage on healthcare expenditure changes.

### Exercise 2

Suppose you want to fit a model so you can use the healthcare expenditure as a percentage of GDP to predict life expectancy. Would a model of the form

$$\text{life_exp} = \beta_0 + \beta_1 ~ \text{health_pct_gdp} + \epsilon$$

be a useful model for the data? Why or why not?
