---
title: "AE 07: Exam 01 review"
subtitle: "Trail users"
date: "February 13, 2025"
editor: visual
---

::: callout-important
Go to the [course GitHub organization](https://github.com/sta210-sp25) and locate your `ae-07`- to get started.

Render, commit, and push your responses to GitHub by the end of class.
:::

## Packages

```{r load-packages}
#| message: false
library(tidyverse)
library(tidymodels)
library(knitr)
```

## Trail users

The Pioneer Valley Planning Commission (PVPC) collected data for ninety days from April 5, 2005 to November 15, 2005. Data collectors set up a laser sensor, with breaks in the laser beam recording when a rail-trail user passed the data collection station.

We will use regression analysis to predict the number of trail users based on weather and other features describing the day.

The variables we'll focus on for this analysis are

-   `volume` estimated number of trail users that day (number of breaks recorded)
-   `hightemp` daily high temperature (in degrees Fahrenheit)
-   `season` one of "Fall", "Spring", or "Summer"
-   `daytype` one of "weekday" or "weekend"

View the data set[^1] to see the remaining variables.

[^1]: Source: [Pioneer Valley Planning Commission](http://www.fvgreenway.org/pdfs/Northampton-Bikepath-Volume-Counts%20_05_LTA.pdf) via the **mosaicData** package.

```{r}
#| message: false

rail_trail <- read_csv("data/rail-trail.csv")
```

## Exploratory analysis

### Exercise 1

Visualize, summarize, and describe the distribution of `volume`.

### Exercise 2

-   Visualize and describe the relationship between `hightemp` and `volume`.
-   Modify the plot to consider if the relationship between these variables differs by `daytype`.

## Modeling

Fit a model using `hightemp` and `daytype` to predict the `volume` for this trail.

### Exercise 3

-   Write the statistical model.

-   Fit the model and write the estimated regression equation. Neatly display the results using 3 digits and the 90% confidence interval for the coefficients.

### Exercise 4

Interpret the slope of `hightemp` in the context of the data.

### Exercise 5

-   Does it make sense to interpret the intercept? Explain your reasoning.
-   If not, what can we do to make the interpretation meaningful?

## Inference for coefficients

### Exercise 6

The following code can be used to create a bootstrap distribution for the model coefficients. Describe what each line of code does, supplemented by any visualizations that might help with your description.

```{r}
#| label: ex-6

set.seed(1234)

boot_dist <- rail_trail |> # <1>
  specify(volume ~ hightemp + daytype) |> # <2>
  generate(reps = 100, type = "bootstrap") |> # <3>
  fit() # <4>
```

1.  \_\_\_

2.  \_\_\_

3.  \_\_\_

4.  \_\_\_

### Exercise 7

Use the bootstrap distribution created in Exercise 6, `boot_dist`, to construct a 90% confidence interval for the coefficient of `hightemp` using bootstrapping and the percentile method and interpret it in context of the data.

### Exercise 8

Conduct a hypothesis test for the coefficient of `hightemp` significance level using permutation with 100 reps. State the hypotheses in words and mathematical notation. Also include a visualization of the null distribution of the slope with the observed slope marked as a vertical line.

### Exercise 9

Now repeat Exercises 7 and 8 using approaches based on mathematical models. You can reference output from previous exercises and/or write new code as needed.

## Inference for prediction

### Exercise 10

Based on your model, predict the volume for a weekday with high temperature of degrees.

### Exercise 11

Suppose you're asked to construct a confidence and a prediction interval for your finding in the previous exercise. Which one would you expect to be wider and why? In your answer clearly state the difference between these intervals.

### Exercise 12

Now construct the intervals and comment on whether your guess is confirmed.

## Interaction terms

### Exercise 13

Now fit the model using `hightemp` and `daytype` to predict volume such that the effect of `hightemp` can differ by `daytype`.

### Exercise 14

-   Write the estimated regression equation for weekends.
-   Write the estimated regression equation for weekdays.

### Exercise 15

According to this model, does the effect of `hightemp` differ for weekends vs. weekdays? Explain.

## Model comparison

### Exercise 16

Which model is a better fit for the data - the model with or without the interaction? Show any work to support your choice.

::: callout-important
To submit the AE:

-   Render the document to produce the PDF with all of your work from today's class.
-   Push all your work to your `ae-07-` repo on GitHub. (You do not submit AEs on Gradescope).
:::
