---
title: "AE 05: Permutation test for the slope"
subtitle: "Houses in Duke Forest"
date: "January 22, 2026"
editor: visual
---

::: callout-important
Go to the [course GitHub organization](https://github.com/sta210-sp26) and locate your `ae-05` repo to get started.
:::

```{r}
#| label: load-packages
#| message: false
#| warning: false

library(tidyverse)
library(tidymodels)
library(openintro)
library(knitr)
```

## Data

The data are on houses that were sold in the Duke Forest neighborhood of Durham, NC around November 2020. It was originally scraped from Zillow, and can be found in the `duke_forest` data set in the **openintro** R package.

**Goal: Use statistical inference to evaluate whether there is a relationship between the age of the house at time of sale and its price.**

## Exploratory data analysis

Let's begin by creating a new variable that is the age of the house in 2020.

```{r}
#| label: create-age-var
duke_forest <- duke_forest |>
  mutate(age_2020 = 2020 - year_built)
```

Now let's visualize the relationship between the age of the house in 2020 and the sales price.

```{r scatterplot}
ggplot(duke_forest, aes(x = age_2020, y = price)) +
  geom_point(alpha = 0.7) +
  labs(
    x = "Age in 2020 (years)",
    y = "Sale price (USD)",
    title = "Price and age of houses in Duke Forest"
  ) +
  scale_y_continuous(labels = label_dollar()) 
```

## Model

```{r fit-model}
df_fit <- lm(price ~ age_2020, data = duke_forest)

tidy(df_fit) |>
  kable(digits = 3)
```

## Hypothesis test

::: callout-tip
For code chunks with fill-in-the-blank code, change code chunk option to `#| eval: true` once you've filled in the code.
:::

### State the null and alternative hypotheses

Write the null and alternative hypotheses in words and mathematical notation.

### Generate null distribution using permutation

Fill in the code, then set `eval: true` .

```{r permutation}
#| eval: false

n = 100
set.seed(012226)

null_dist <- _____ |>
  specify(______) |>
  hypothesize(null = "independence") |>
  generate(reps = _____, type = "permute") |>
  fit()
```

### Visualize distribution

```{r null-dist}
# Code for histogram of null distribution
```

### Compute the p-value

```{r calc-pvalue}
#| eval: false 

# get observed fit 
observed_fit <- duke_forest |>
  specify(price ~ age_2020) |>
  fit()

# calculate p-value
get_p_value(
  ____,
  obs_stat = ____,
  direction = "two-sided"
)
```

### State conclusion

Write your conclusion in the context of the data. You can use 0.05 as the decision-making threshold.

## Bootstrap CI

### Construct the bootstrap CI

Construct a 95% bootstrap confidence interval.

```{r}
#| label: construct-ci

# Code for ci

```

### Draw conclusion

-   Interpret the interval in the context of the data.

-   Is the interval consistent with the conclusion from your hypothesis test? Briefly explain why or why not.

# Wrapping up

::: callout-important
Once you’ve completed the AE:

-   Render the document to produce the PDF with all of your work from today’s class.

-   Push all your work to your AE repo on GitHub. You’re done! 🎉
:::
