---
title: "AE 06: Model comparison"
subtitle: "Restuarant tips"
date: "February 6, 2025"
editor: visual
---

::: callout-important
Render, commit, and push your responses to GitHub by the end of class.
:::

```{r}
#| label: load-packages
#| warning: false
#| message: false
library(tidyverse)
library(tidymodels)
library(knitr)
```

## Data

Which variables help us predict the amount customers tip at a restaurant? To answer this question, we will use data collected in 2011 by a student at St. Olaf who worked at a local restaurant.

The variables we’ll focus on for this analysis are

-   `Tip`: amount of the tip

-   `Party`: number of people in the party

-   `Meal`: Time of day (Lunch, Dinner, Late Night)

-   `Age`: Age category of person paying the bill (Yadult, Middle, SenCit)

-   `Day`: Day of the week (includes every day but Monday)

View the data set to see the remaining variables.

```{r}
#| label: load-data
#| warning: false
#| message: false
tips <- read_csv("data/tip-data.csv")
```

## Exercise 1

Split the data into training (80%) and testing (20%) sets. Use seed `2025`.

## Exercise 2

Use the training data to fit a model using `Party`, `Age`, and `Meal` to predict tips. Compute the $R^2$ and $Adj. R^2$ for this model.

## Exercise 3

Now fit a model predicting tips using `Party`, `Age`, and `Meal`, such that the effect of party can differ by `Meal`. Compute $R^2$ and $Adj. R^2$ for this model.

## Exercise 4

-   Which model do you choose - the model from Exercise 2 or Exercise 3? Why?

-   Compute RMSE for the selected model.

## Exercise 5

Now let's use the testing data to assess the performance of the model selected in Exercise 4.

-   Compute the predicted tips for the testing data. Add the predictions to the testing data set.

-   Compute RMSE and $R^2$ for the testing data.

```{r}
#| label: ex-5

# add code here

```

## Exercise 6

-   How do RMSE compare between the training and testing data?

-   How does $R^2$ compare between the training and testing data?

-   Is this what you expect? Why?

## Exercise 7

Why can we use $R^2$ as an assessment of performance on the testing data even if we can't use it to compare models?

## To submit the AE:

::: callout-important
-   Render the document to produce the PDF with all of your work from today's class.
-   Push all your work to your repo on GitHub. (You do not submit AEs on Gradescope).
:::
